---
title: "Descriptive Statistics"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(dplyr)
library(ggplot2)
library(readr)
library(janitor)
library(stringr)
library(tidyverse)
library(expss)
#library(qwraps2)
#options(qwraps2_markup = "markdown")

```

```{r read data, include = FALSE}

data = 
  read_csv("Data/SearchResultsTable.csv") %>%
  clean_names() 

gantt_data = readRDS("Data/gantt_data.rds") %>%
   mutate(date_start2 = format(date_start, format = "%B %d, %Y") )

# Create dataset for descriptives for clinical trials in the US
us_studies = data %>%
  # Filter only studies that have sites in the United States
  filter(str_detect(locations, "United States"))  %>%
         # Count number of locations, interventions, and funders
  mutate(num_locs = as.numeric(str_count(locations, "\\|")) + 1,
         num_intervs = as.numeric(str_count(interventions, "\\|")) + 1,
         num_funders = as.numeric(str_count(funded_bys, "\\|")) + 1,
         # Dummy for Funded by - "Industry"
         fundby_industry = ifelse(str_detect(funded_bys, "Industry") == TRUE, 1, 0),
         # dummy for Funded by - "NIH"
         fundby_nih = ifelse(str_detect(funded_bys, "NIH") == TRUE, 1, 0),
         # dummy for FUnded by - "U.S. Fed"
         fundby_usfed = ifelse(str_detect(funded_bys, "U.S. Fed") == TRUE, 1, 0),
         # dummy for Funded by - "Other")
         fundby_other = ifelse(str_detect(funded_bys, "Other") == TRUE, 1, 0),
         # Intervention Types -- Dummy Variables
         interv_drug =  ifelse(str_detect(interventions, "Drug:") == TRUE, 1, 0), 
         interv_behav = ifelse(str_detect(interventions, "Behavioral:") == TRUE, 1, 0),
         interv_biol  = ifelse(str_detect(interventions, "Biological:") == TRUE, 1, 0), 
         interv_diag  = ifelse(str_detect(interventions, "Diagnostic Test:") == TRUE, 1, 0) ,
         interv_device = ifelse(str_detect(interventions, "Device:") == TRUE, 1, 0),
         interv_rad   = ifelse(str_detect(interventions, "Radiation:") == TRUE, 1, 0),
         interv_diet  = ifelse(str_detect(interventions, "Dietary Supplement:") == TRUE, 1, 0),
         interv_gene  = ifelse(str_detect(interventions, "Genetic:") == TRUE, 1, 0),
         interv_proc  = ifelse(str_detect(interventions, "Procedure:") == TRUE, 1, 0),
         interv_other = ifelse(str_detect(interventions, "Other:") == TRUE, 1, 0),
         # Enrollment number as numeric
         enrol_num = as.numeric(enrollment),
         mm_start   = str_sub(start_date,  1,  3), # Extract month
         yy_start   = str_sub(start_date, -4, -1), # Extract 4-digit year
         date_start = as.Date(paste(yy_start, mm_start, "01", sep = "-"), "%Y-%b-%d"),
         date_start2 = format(date_start, "%B %d, %Y"))

# Collapse study type and phase into one variable
us_studies$type_phase = NA

us_studies$type_phase[str_detect(us_studies$phases, "Phase 1") == TRUE] <- "Phase 1/Phase 2"
us_studies$type_phase[str_detect(us_studies$phases, "Phase 2") == TRUE] <- "Phase 1/Phase 2"
us_studies$type_phase[str_detect(us_studies$phases, "Phase 3") == TRUE] <- "Phase 3/Phase 4"
us_studies$type_phase[str_detect(us_studies$phases, "Phase 4") == TRUE] <- "Phase 3/Phase 4"
us_studies$type_phase[is.na(us_studies$phases)] <- "Unknown Phase"
us_studies$type_phase[us_studies$study_type == "Observational"] <- "Observational"




```


The primary datafile (`SearchResultsTable.csv`) contained informaton on **`r format(nrow(data), nsmall = 1, big.mark = ",")` Alzheimer's disease studies** registered with [ClinicalTrials.gov](http://clinicaltrials.gov).  We restricted our analysis on the **`r nrow(us_studies)` studies with study sites in the United States**, and as of the November 9, 2017 data pull, the earliest clinical trial on Alzheimer's Disease started on  `r  us_studies %>% pull(date_start2) %>% min(na.rm=TRUE)` and the most recent on `r  us_studies %>% pull(date_start2) %>% max(na.rm=TRUE)`.

We categorized the trials into 5 main categories based on study type and phase: 

* Observational: if a study is of study type observational **(n = `r us_studies %>% filter(type_phase == "Observational") %>% nrow()`)**

* Phase 1/2: if a study is interventional, and phase is classifed as early Phase 1, Phase 1, or combined Phase 1/Phase 2 **(n = `r us_studies %>% filter(type_phase == "Phase 1/Phase 2") %>% nrow()`)**

* Phase 3/4: if a study is interventional, and phase is classified as Phase 3, Phase 4, or combined Phase 3/Phase 4 **(n = `r us_studies %>% filter(type_phase == "Phase 3/Phase 4") %>% nrow()`)**

* Unknown Phase: if a study is interventional, and phase was not specified **(n = `r us_studies %>% filter(type_phase == "Unknown Phase") %>% nrow()`)**

For each of the five categories, we examined the number of study sites, number of funders, and number of interventions (see Table X below). 


```{r Descriptives table, echo = FALSE, eval=TRUE}
# Add variable labels
us_studies = apply_labels(us_studies,
  type_phase = "Study Type",
  num_locs = "No. Study Sites",
  num_funders = "No. Funders",
  num_intervs = "No. Interventions")

# Create summary table
us_studies %>%
  tab_cells(num_locs) %>%
  tab_cols(type_phase) %>%
  tab_stat_mean() %>%
  tab_stat_sd() %>%
  tab_stat_median() %>%
  tab_stat_min() %>%
  tab_stat_max() %>%
  
  tab_cells( num_funders) %>%
  tab_cols(type_phase) %>%
  tab_stat_mean() %>%
  tab_stat_sd() %>%
  tab_stat_median() %>%
  tab_stat_min() %>%
  tab_stat_max() %>%
  
  tab_cells(num_intervs) %>%
  tab_cols(type_phase) %>%
  tab_stat_mean() %>%
  tab_stat_sd() %>%
  tab_stat_median() %>%
  tab_stat_min() %>%
  tab_stat_max() %>%
  tab_pivot() %>%
  htmlTable(caption = "Descriptive statistics by study type")


```

We also examined the number of new studies each year.

```{r number of studies by year, echo=FALSE, warning=FALSE}
us_studies %>%
  mutate(year_start   = as.numeric(str_sub(start_date, -4, -1))) %>%
  filter(!is.na(year_start) & year_start <= 2017) %>%
  group_by(year_start) %>%
  count() %>%
  ggplot(aes(x = year_start, y = n)) + 
    geom_point() +
    geom_line() +
    theme_bw() +
    labs(x = "Study start (year)", 
         y = "No. of studies", 
         title = "No. of new Alzheimer's Disease studies per year") +
    scale_x_continuous(breaks = seq(1980, 2017, 5))

us_studies %>%
  mutate(year_start   = as.numeric(str_sub(start_date, -4, -1))) %>%
  filter(!is.na(year_start) & year_start <= 2017) %>%
  group_by(type_phase,year_start) %>%
  count() %>%
  ggplot(aes(x = year_start, y = n, color = type_phase)) + 
    geom_point() +
    geom_line() +
    theme_bw() +
    labs(x = "Study start (year)", y = "No. of studies", 
         title = "No. of new Alzheimer's Disease studies per year by study type",
         color = "Study type") +
    scale_x_continuous(breaks = seq(1980, 2017, 5)) +
    theme(legend.position = "bottom")

```



